---
title: "Project Metadata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project Metadata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
In this vignette we will ingest project metadata from a CSV, restructure the data, and then create a json object that can be validated.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wddsWizard)
```

## Load in the CSV and clean it up

Our example data requires some light cleaning to make transforming it into csv
easier. 

```{r load-data}
project_metadata <- read.csv("../inst/extdata/example_data/example_project_metadata .csv")

## turn empty strings into NAs in the group field
project_metadata <- project_metadata |>
  dplyr::mutate(Group = dplyr::case_when(
    Group != "" ~ Group,
    TRUE ~ NA
  )) 

## use `fill` to complete the items column and `mutate` to make groups a little
## more ergonomic

 project_metadata_filled <- tidyr::fill(data = project_metadata,Group) 
```

## Restructure data

The validation schema is expecting JSON, so we have to restructure the data into
a list that can be converted to JSON.

For Creators, Resources, and Funding References, its possible to have multiple 
entities in each group. In our example data, there are two creators and three funding references. So we need to pull out the `entity_id`s for the creators and funding references then clean up the `Group` field so it can be used a general category
for Creators, Resources, and Funding References.

```{r restructure}

# get ids for components of a group.
project_metadata_ids <- project_metadata_filled |>
  dplyr::mutate(
    entity_id = stringr::str_extract(string = Group,pattern = "[0-9]"),
    # make sure that there are no NA entity IDs
    entity_id = dplyr::case_when(
      is.na(entity_id) ~ "1",
      TRUE ~ entity_id
    )
  ) |>
  # drop entity ids from group field and convert to camel case
   dplyr::mutate(Group = stringr::str_replace_all(string = Group,pattern = " [0-9]", replacement = ""),
                 Group = snakecase::to_lower_camel_case(Group))

 ## split dataframe by Group for further processing
 
project_metadata_list  <- split(project_metadata_ids,project_metadata_ids$Group)


# this function creates entity json

get_entity <- function(x){
  y <- x |>
    dplyr::select(-Group,-entity_id) |>
    dplyr::filter(Value != "") |>
    dplyr::mutate(Variable = snakecase::to_lower_camel_case(Variable))
  
  z <- tidyr::pivot_wider(y,names_from = Variable, values_from = Value) 
    
  
  return(z)
}

project_metadata_list_entities <- purrr::map(project_metadata_list,function(x){
  if(all(x$entity_id == "1")){
    out <- get_entity(x)
    return(out)
  }
  
  x_list <- split(x,x$entity_id)
  names(x_list) <- NULL
  out <-purrr::map(x_list, get_entity)
  return(out)
})

jsonlite::toJSON(project_metadata_list_entities,pretty = TRUE,dataframe = "columns")

```

### Make Creator Objects
Now that we have our groups in their own data frames, we can translate them into
json. 


```{r creators}
jsonlite::toJSON(project_metadata_list,pretty = TRUE,dataframe = "columns")

```
