---
title: "Schema Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Schema Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document provides background information about json-schemas in general and
more information about the structure of the wildlife disease data standard. 

## Background

### Why do we need a data standard?

Data standards facilitate the sharing, reuse, and aggregation of data by
humans and machines through the use of a common structure and vocabulary. When
a dataset meets a data standard, it can be combined with other data sets that 
meet that standard. As the number of data sets that use the standard grows,
the questions that researchers can ask of the data changes. For example, if we have
data sets describing disease incidence for some pathogen from thousands of 
locations across multiple regions, we can explore general modeling frameworks for
that pathogen. Being able to reuse and aggregate data sets is especially 
important in fields like disease ecology where studies are time, resource, and 
labor intensive. 

### What is a JSON-schema

A JSON-schema is a human and machine readable document that
prescribes the structure, constraints, and content of a data set that meets a particular standard.
By constraints and contents, we mean rules around the type of data used in  particular field
(character, numeric, logical, etc), and its values (e.g. `massUnits` must be one 
of kg, mg, or g; `latitude` must be between -90 and 90; `sampleID` must be unique). 

Schema properties may themselves be schema. For example, the `data` property in this
standard defines a `data` object that is a flat table with constraints, types, and/or
requirements. In this way, JSON-schema allows for the construction of modular 
schema documents. 

Once we have created a schema, we can then validate data against it. The validation
process happens via a validation engine and tells us if the data conform
to the standard. If the data do not conform, then the validation engine tells
us precisely where the data are non-conformant and what the data standard expected
to see. 

![](https://json-schema.org/img/json_schema.svg){width=95%}

For more detailed information see [JSON-Schema.org](https://json-schema.org/overview/what-is-jsonschema)

## How do I use it?

### Data
#### Starting from scratch

Use our CSV templates and the data standard as a guide for storing data. This may not be how data are
initially stored after collection, but make sure that at least the minimum required
fields are captured in data collection and that they match the data standard. 

#### Starting from an existing project



This package has vignettes describing how to move from CSV/excel to JSON for
[project metadata](project_metadata.html) and [wildlife disease data](disease_data.html).





## Wildlife Disease Data Standard (WDDS) Structure

The Wildlife Disease Data Standard has eight required fields:`data`, `methodology`, `creators`, `titles`, `publicationYear`, `descriptions`, `language`, and `fundingReferences`. These
fields must present



### Schema terms

```{r schema-as-table}

library(kableExtra)
wdds_terms <- read.csv(here::here("inst/extdata/example_data/wdds_terms.csv"))



wdds_terms |>
  kableExtra::kbl() |>
  kableExtra::kable_paper(lightable_options = "striped",full_width = F) |>
  kableExtra::column_spec(2) |>
  kableExtra::column_spec(3, width = "30em")


```



## Data Package Structure

## Other packages of interest
- deposits
- jsonvalidate

