---
title: "Schema Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Schema Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document provides background information about json-schemas in general and
more information about the structure of the Wildlife Disease Data standard. 

## Background

### Why do we need a data standard?

Data standards facilitate the sharing, reuse, and aggregation of data by
humans and machines through the use of a common structure and vocabulary. When
a dataset meets a data standard, it can be combined with other data sets that 
meet that standard. As the number of data sets that use the standard grows,
the questions that researchers can ask of the data changes. For example, if we have
data sets describing disease incidence for some pathogen from thousands of 
locations across multiple regions, we can explore general modeling frameworks for
that pathogen. Being able to reuse and aggregate data sets is especially 
important in fields like disease ecology where studies are time, resource, and 
labor intensive. 

### What is a JSON-schema

A JSON-schema is a human and machine readable document that
prescribes the structure, constraints, and content of a data set that meets a particular standard.
By constraints and contents, we mean rules around the type of data used in  particular field
(character, numeric, logical, etc), and its values (e.g. `massUnits` must be one 
of kg, mg, or g; `latitude` must be between -90 and 90; `sampleID` must be unique). 

Schema properties may themselves be schema. For example, the `data` property in this
standard defines a `data` object that is a flat table with constraints, types, and/or
requirements. In this way, JSON-schema allows for the construction of modular 
schema documents. 

Once we have created a schema, we can then validate data against it. The validation
process happens via a validation engine and tells us if the data conform
to the standard. If the data do not conform, then the validation engine tells
us precisely where the data are non-conformant and what the data standard expected
to see. 

![](https://json-schema.org/img/json_schema.svg){width=95%}

For more detailed information see [JSON-Schema.org](https://json-schema.org/overview/what-is-jsonschema)

## How do I use it this package and the data standard?

#### Starting from scratch

Use our CSV templates and the data standard as a guide for storing data. Make sure
that at least the minimum required fields are captured in data collection,
and that they match the data standard. Lab and field data may not immediately
come out in a "tidy" format but as long as all properties are captured it 
should be possible to reshape the data to be compatible with the standard.

#### Starting from an existing project

This is more challenging than starting from scratch and requires harmonization 
your data and the data standard. Once you have crosswalked the fields from your
data onto the standard, you can start to reshape it into a single, tidy, table.
This package has vignettes describing how to move from CSV/excel to JSON for
[project metadata](project_metadata.html) and [wildlife disease data](disease_data.html).


## Wildlife Disease Data Standard (WDDS) Structure

The Wildlife Disease Data Standard is composed of two sub-schemas (1) `data` and
(2) `project_metadata`. 

`data` describes the structure and contents of the wildlife disease data. 
It has certain required fields and is extensible. This data should be stored as
[tidy](https://doi.org/10.18637/jss.v059.i10) dataset in a flat file like a CSV.
This component of the standard relies heavily on the Darwin Core data standard.

`project_metadata` describes the structure and contents of the descriptive 
metadata. That is, metadata about the project that enables discovery, 
identification, and attribution. This component of the standard relies heavily
on the Data Cite Metadata Schema.

Researchers may validate their data against each sub-schema separately, or
use them in tandem to validate an entire data package. The term "data package" 
refers to a list or JSON object that contains both the `data` and 
`project_metadata` components.

"array items" - these can be thought of as vectors or columns in a data frame. These will be of one type and may allow for null values.  

### Schema terms

```{r terms, eval=FALSE, include=FALSE, results="asis"}
# library(wddsWizard)
create_schema_docs() |>
  cat()
```
## Other packages of interest
- deposits
- jsonvalidate
## data  
**Description**: **REQUIRED** Wildlife disease data. Stored in tidy form.  
 **Reference**: schemas/disease_data.json  

 ### sampleID  
 **Type**: array  
 **Description**: **REQUIRED** A researcher-generated unique ID for the sample: usually a unique string of both characters and integers (e.g., OS BZ19-114 to indicate an oral swab taken from animal BZ19-114; see worked example below), to avoid conflicts that can arise when datasets are merged with number-only notation for samples. Ideally, sample names should be kept consistent across all online databases and physical resources (e.g., museum collections or project-specific sample archives). <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### animalID  
 **Type**: array  
 **Description**: A researcher-generated unique ID for the individual animal from which the sample was collected: usually a unique string of both characters and integers (e.g., BZ19-114 to indicate animal 114 sampled in 2019 in Belize). Ideally, animal names should again be kept consistent across online databases and physical resources. <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### latitude  
 **Type**: array  
 **Description**: **REQUIRED** Latitude of the collection site in decimal format. See http://rs.tdwg.org/dwc/terms/decimalLatitude <details><summary> Array Items </summary>
 - **type**: number, null  
- **minItems**: 1  
- **maximum**: 90  
- **minimum**: -90   </details>  

### longitude  
 **Type**: array  
 **Description**: **REQUIRED** Longitude of the collection site  in decimal format. See http://rs.tdwg.org/dwc/terms/decimalLongitude <details><summary> Array Items </summary>
 - **type**: number, null  
- **minItems**: 1  
- **maximum**: 180  
- **minimum**: -180   </details>  

### spatialUncertaintiy  
 **Type**: array  
 **Description**: Coordinate uncertainty from GPS recordings, post-hoc digitization, or systematic alterations (e.g., jittering or rounding) expressed in meters. See http://rs.tdwg.org/dwc/terms/coordinateUncertaintyInMeters <details><summary> Array Items </summary>
 - **type**: number, null  
- **minItems**: 1  
- **minimum**: 0   </details>  

### collectionDay  
 **Type**: array  
 **Description**: The day of the month on which the specimen was collected. See http://rs.tdwg.org/dwc/terms/day <details><summary> Array Items </summary>
 - **type**: integer, null  
- **minItems**: 1  
- **minimum**: 1  
- **maximum**: 31   </details>  

### collectionMonth  
 **Type**: array  
 **Description**: The month in which the specimen was collected. See http://rs.tdwg.org/dwc/terms/month <details><summary> Array Items </summary>
 - **type**: integer, null  
- **minItems**: 1  
- **minimum**: 1  
- **maximum**: 12   </details>  

### collectionYear  
 **Type**: array  
 **Description**: The year in which the specimen was collected. See http://rs.tdwg.org/dwc/terms/year <details><summary> Array Items </summary>
 - **type**: integer, null   </details>  

### sampleCollectionMethod  
 **Type**: array  
 **Description**: **REQUIRED** The technique used to acquire the sample and/or the tissue from which the sample was acquired (e.g. visual inspection; swab; wing punch; necropsy). <details><summary> Array Items </summary>
 - **type**: string  
- **minItems**: 1   </details>  

### sampleMaterial  
 **Type**: array  
 **Description**: Organic material from the animal body (i.e., tissue or fluid) extracted into a sample (e.g., “liver”; “blood”; “skin”; “whole organism”). <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### sampleCollectionBodyPart  
 **Type**: array  
 **Description**: Part of the animal body that samples are generated or collected from (e.g., “rectum”; “wing”). <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### hostIdentification  
 **Type**: array  
 **Description**: **REQUIRED** The Linnaean classification of the animal from which the sample was collected, reported at the lowest possible level (ideally, species binomial name: e.g., Odocoileus virginianus or Ixodes scapularis). As necessary, researchers may also include an additional field indicating when uncertainty exists in the identification of the host organism (see Adding new fields). See http://rs.tdwg.org/dwc/terms/scientificName <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1  
- **not**: [HOMOhomo]{4} [SAPIENSsapiens]{7}   </details>  

### organismSex  
 **Type**: array  
 **Description**: The sex of the individual animal from which the sample was collected. See http://rs.tdwg.org/dwc/terms/sex <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### liveCapture  
 **Type**: array  
 **Description**: Whether the individual animal from which the sample was collected was alive at the time of capture. Should be TRUE or FALSE; lethal sampling should be recorded as TRUE as this field describes the organism at the time of capture. <details><summary> Array Items </summary>
 - **type**: boolean, null  
- **minItems**: 1   </details>  

### healthNotes  
 **Type**: array  
 **Description**: Any additional (unstructured) notes about the state of the animal, such as disease presentation. <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### hostLifeStage  
 **Type**: array  
 **Description**: The life stage of the animal from which the sample was collected (as appropriate for the organism) (e.g., juvenile, adult). See http://rs.tdwg.org/dwc/terms/lifeStage <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### age  
 **Type**: array  
 **Description**: The numeric age of the animal from which the sample was collected, at the time of sample collection, if known (e.g., in monitored populations). <details><summary> Array Items </summary>
 - **type**: number, null  
- **minItems**: 1  
- **minimum**: 0   </details>  

### ageUnits  
 **Type**: array  
 **Description**: The units in which age is measured (usually years). <details><summary> Array Items </summary>
 - **type**: string, null  
- **enum**: years, months, days, hours, minutes, seconds  
- **minItems**: 1   </details>  

### mass  
 **Type**: array  
 **Description**: The mass of the animal from which the sample was collected, at the time of sample collection. <details><summary> Array Items </summary>
 - **type**: number, null  
- **minItems**: 1  
- **minimum**: 0   </details>  

### massUnits  
 **Type**: array  
 **Description**: The units that mass is recorded in (e.g., kg). <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### length  
 **Type**: array  
 **Description**: The numeric length of the animal from which the sample was collected, at the time of sample collection. <details><summary> Array Items </summary>
 - **type**: number, null  
- **minItems**: 1  
- **minimum**: 0   </details>  

### lengthMeasurement  
 **Type**: array  
 **Description**: The axis of measurement for the organism being measured (e.g., snout-vent length or just SVL; wing length; primary feather). <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### lengthUnits  
 **Type**: array  
 **Description**: The units that length is recorded in (e.g., meters). <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### organismQuantity  
 **Type**: array  
 **Description**: A number or enumeration value for the quantity of organisms. See http://rs.tdwg.org/dwc/terms/organismQuantity <details><summary> Array Items </summary>
 - **type**: number, null  
- **minItems**: 1  
- **minimum**: 0   </details>  

### organismQuantityUnits  
 **Type**: array  
 **Description**: The units that organism quantity is recorded in (e.g. “individuals”). See http://rs.tdwg.org/dwc/iri/organismQuantityType <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### detectionTarget  
 **Type**: array  
 **Description**: **REQUIRED** The taxonomic identity of the parasite being screened for in the sample. This will often be coarser than the identity of a specific parasite identified in the sample: for example, in a study screening for novel bat coronaviruses, the entire family Coronaviridae might be the target; in a parasite dissection, the targets might be Acanthocephala, Cestoda, Nematoda, and Trematoda. For deep sequencing approaches (e.g., metagenomic and metatranscriptomic viral discovery), researchers should report each alignment target used as a new test to maximize reporting of negative data, or alternatively, select a subset that reflect specific study objectives and the focus of analysis (e.g., specific viral families). See http://rs.tdwg.org/dwc/terms/associatedOccurrences <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### detectionMethod  
 **Type**: array  
 **Description**: **REQUIRED** The type of test performed to detect the parasite or parasite-specific antibody (e.g., 'qPCR', ‘ELISA’) <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### forwardPrimerSequence  
 **Type**: array  
 **Description**: The sequence of the forward primer used for parasite detection (e.g., for a pan-coronavirus primer: 5’ CDCAYGARTTYTGYTCNCARC 3'). (Strongly encouraged if applicable, e.g., for PCR.) <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### reversePrimerSequence  
 **Type**: array  
 **Description**: The sequence of the reverse primer used for parasite detection (e.g., 5’ RHGGRTANGCRTCWATDGC 3'). (Strongly encouraged if applicable, e.g., for PCR.) <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### geneTarget  
 **Type**: array  
 **Description**: The parasite gene targeted by the primer (e.g. “RdRp” for PCR.). <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### primerCitation  
 **Type**: array  
 **Description**: Citation(s) for the primer(s) (ideally doi, or other permanent identifier for a work, e.g. PMID).  <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### probeTarget  
 **Type**: array  
 **Description**: Antibody or antigen targeted for detection. (Strongly encouraged if applicable, e.g., for ELISA.) <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### probeType  
 **Type**: array  
 **Description**: Antibody or antigen used for detection. (Strongly encouraged if applicable, e.g., for ELISA.) <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### probeCitation  
 **Type**: array  
 **Description**: Citation(s) for the probe(s) (ideally doi, or other permanent identifier for a work, e.g. PMID).  <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### detectionOutcome  
 **Type**: array  
 **Description**: **REQUIRED** The test result (i.e., positive, negative, or inconclusive). To avoid ambiguity, these specific values are suggested over numeric values (0 or 1). See http://rs.tdwg.org/dwc/terms/occurrenceStatus <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### detectionMeasurement  
 **Type**: array  
 **Description**: Any numeric measurement of parasite detection that is more detailed than simple positive or negative results (e.g., viral titer, parasite counts, sequence reads). <details><summary> Array Items </summary>
 - **type**: number, null  
- **minItems**: 1   </details>  

### detectionMeasurementUnits  
 **Type**: array  
 **Description**: Units for quantitative measurements of parasite intensity or test results (e.g., Ct, TCID50/mL, or parasite count). <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### parasiteIdentification  
 **Type**: array  
 **Description**: **REQUIRED** The identity of a parasite detected by the test, if any, reported to the lowest possible taxonomic level, either as a Linnaean binomial classification or within the convention of a relevant taxonomic authority (e.g., Borrelia burgdorferi or Zika virus). Parasite identification may be more specific than detection target. <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### parasiteID  
 **Type**: array  
 **Description**: A researcher-generated unique ID for an individual parasite (primarily useful in nested cases where this ID is used as an animal ID in another row, such as pathogen  testing of a blood-feeding arthropod removed from a vertebrate host). <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### parasiteLifeStage  
 **Type**: array  
 **Description**: The life stage of the parasite from which the sample was collected (as appropriate for the organism) (e.g., juvenile, adult). <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  

### genbankAccession  
 **Type**: array  
 **Description**: The GenBank accession for any parasite genetic sequence(s), if appropriate.  Accession numbers or other identifiers for related data stored on another platform should be added in a different field (e.g. GISAID Accession, Immport Accession). See http://rs.tdwg.org/dwc/terms/otherCatalogNumbers  <details><summary> Array Items </summary>
 - **type**: string, null  
- **minItems**: 1   </details>  
 

## project_metadata  
**Description**: **REQUIRED** Metadata for a project that largely follows the Datacite data standard.  
 **Reference**: schemas/project_metadata.json  

 ### methodology  
 **Type**: object  
 **Description**: **REQUIRED** broad categorization of how data were collected.  

### creators  
 **Type**: array  
 **Description**: **REQUIRED** The full names of the creators. Should be in the format familyName, givenName. <details><summary> Array Items </summary>
 - ### name  
 **Type**: string  
 **Description**: **REQUIRED** [DataCite name](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=name&check_keywords=yes&area=default)  

- ### nameType  
 **Type**: string  
 **Description**: [DataCite nameType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=nameType&check_keywords=yes&area=default)  

- ### givenName  
 **Type**: string  
 **Description**: [DataCite givenName](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=givenName&check_keywords=yes&area=default)  

- ### familyName  
 **Type**: string  
 **Description**: [DataCite familyName](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=familyName&check_keywords=yes&area=default)  

- ### nameIdentifiers  
 **Type**: array  
 **Description**: [DataCite nameIdentifiers](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=nameIdentifiers&check_keywords=yes&area=default) <details><summary> Array Items </summary>
 - [DataCite nameIdentifiers](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=nameIdentifiers&check_keywords=yes&area=default) <details><summary> Array Items </summary>
 - ### nameIdentifier  
 **Type**: string  
 **Description**: **REQUIRED** [DataCite nameIdentifier](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=nameIdentifier&check_keywords=yes&area=default)  

- ### nameIdentifierScheme  
 **Type**: string  
 **Description**: **REQUIRED** [DataCite nameIdentifierScheme](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=nameIdentifierScheme&check_keywords=yes&area=default)  

- ### schemeUri  
 **Type**: string  
 **Description**: [DataCite schemeUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=schemeUri&check_keywords=yes&area=default)  
 </details> </details>  

- ### affiliation  
 **Type**: array  
 **Description**: [DataCite affiliation](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=affiliation&check_keywords=yes&area=default) <details><summary> Array Items </summary>
 - [DataCite affiliation](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=affiliation&check_keywords=yes&area=default) <details><summary> Array Items </summary>
 - ### name  
 **Type**: string  
 **Description**: **REQUIRED** [DataCite name](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=name&check_keywords=yes&area=default)  

- ### affiliationIdentifier  
 **Type**: string  
 **Description**: [DataCite affiliationIdentifier](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=affiliationIdentifier&check_keywords=yes&area=default)  

- ### affiliationIdentifierScheme  
 **Type**: string  
 **Description**: [DataCite affiliationIdentifierScheme](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=affiliationIdentifierScheme&check_keywords=yes&area=default)  

- ### schemeUri  
 **Type**: string  
 **Description**: [DataCite schemeUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=schemeUri&check_keywords=yes&area=default)  
 </details> </details>  

- ### lang  
 **Type**: string  
 **Description**: [DataCite lang](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=lang&check_keywords=yes&area=default)  

- **allOf**: #/definitions/creator  
- **required**: name   </details>  

### titles  
 **Type**: array  
 **Description**: **REQUIRED** A name or title by which a resource is known. <details><summary> Array Items </summary>
 - **REQUIRED** A name or title by which a resource is known. <details><summary> Array Items </summary>
 - ### title  
 **Type**: string  
 **Description**: **REQUIRED** [DataCite title](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=title&check_keywords=yes&area=default)  

- ### titleType  
 **Type**: string  
 **Description**: [DataCite titleType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=titleType&check_keywords=yes&area=default)  

- ### lang  
 **Type**: string  
 **Description**: [DataCite lang](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=lang&check_keywords=yes&area=default)  
 </details> </details>  

### identifier  
 **Type**: array  
 **Description**: A unique string that identifies a resource. <details><summary> Array Items </summary>
 - A unique string that identifies a resource. <details><summary> Array Items </summary>
 - ### identifier  
 **Type**: string  
 **Description**: **REQUIRED** [DataCite identifier](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=identifier&check_keywords=yes&area=default)  

- ### identifierType  
 **Type**: string  
 **Description**: [DataCite identifierType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=identifierType&check_keywords=yes&area=default)  
 </details> </details>  

### publicationYear  
 **Type**: string  
 **Description**: **REQUIRED** The year when the data was or will be made publicly available.  

### rights  
 **Type**: array  
 **Description**: Any rights information for this resource. <details><summary> Array Items </summary>
 - Any rights information for this resource. <details><summary> Array Items </summary>
 - ### rights  
 **Type**: string  
 **Description**: [DataCite rights](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=rights&check_keywords=yes&area=default)  

- ### rightsUri  
 **Type**: string  
 **Description**: [DataCite rightsUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=rightsUri&check_keywords=yes&area=default)  

- ### rightsIdentifier  
 **Type**: string  
 **Description**: [DataCite rightsIdentifier](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=rightsIdentifier&check_keywords=yes&area=default)  

- ### rightsIdentifierScheme  
 **Type**: string  
 **Description**: [DataCite rightsIdentifierScheme](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=rightsIdentifierScheme&check_keywords=yes&area=default)  

- ### schemeUri  
 **Type**: string  
 **Description**: [DataCite schemeUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=schemeUri&check_keywords=yes&area=default)  

- ### lang  
 **Type**: string  
 **Description**: [DataCite lang](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=lang&check_keywords=yes&area=default)  
 </details> </details>  

### descriptions  
 **Type**: array  
 **Description**: **REQUIRED** All additional information that does not fit in any of the other categories. May be used for technical information or detailed information associated with a scientific instrument. <details><summary> Array Items </summary>
 - **REQUIRED** All additional information that does not fit in any of the other categories. May be used for technical information or detailed information associated with a scientific instrument. <details><summary> Array Items </summary>
 - ### description  
 **Type**: string  
 **Description**: **REQUIRED** [DataCite description](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=description&check_keywords=yes&area=default)  

- ### descriptionType  
 **Type**: string  
 **Description**: **REQUIRED** [DataCite descriptionType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=descriptionType&check_keywords=yes&area=default)  

- ### lang  
 **Type**: string  
 **Description**: [DataCite lang](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=lang&check_keywords=yes&area=default)  
 </details> </details>  

### language  
 **Type**: string  
 **Description**: **REQUIRED** The primary language of the resource.  

### fundingReferences  
 **Type**: array  
 **Description**: **REQUIRED** Name and other identifying information of a funding provider. <details><summary> Array Items </summary>
 - **REQUIRED** Name and other identifying information of a funding provider. <details><summary> Array Items </summary>
 - ### funderName  
 **Type**: string  
 **Description**: **REQUIRED** [DataCite funderName](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=funderName&check_keywords=yes&area=default)  

- ### funderIdentifier  
 **Type**: string  
 **Description**: [DataCite funderIdentifier](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=funderIdentifier&check_keywords=yes&area=default)  

- ### funderIdentifierType  
 **Type**: string  
 **Description**: [DataCite funderIdentifierType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=funderIdentifierType&check_keywords=yes&area=default)  

- ### awardNumber  
 **Type**: string  
 **Description**: [DataCite awardNumber](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=awardNumber&check_keywords=yes&area=default)  

- ### awardUri  
 **Type**: string  
 **Description**: [DataCite awardUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=awardUri&check_keywords=yes&area=default)  

- ### awardTitle  
 **Type**: string  
 **Description**: [DataCite awardTitle](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=awardTitle&check_keywords=yes&area=default)  
 </details> </details>  

### relatedIdentifiers  
 **Type**: array  
 **Description**: [DataCite relatedIdentifiers](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=relatedIdentifiers&check_keywords=yes&area=default) <details><summary> Array Items </summary>
 - ### relationType  
 **Type**: string  
 **Description**: **REQUIRED** [DataCite relationType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=relationType&check_keywords=yes&area=default)  

- ### relatedMetadataScheme  
 **Type**: string  
 **Description**: [DataCite relatedMetadataScheme](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=relatedMetadataScheme&check_keywords=yes&area=default)  

- ### schemeUri  
 **Type**: string  
 **Description**: [DataCite schemeUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=schemeUri&check_keywords=yes&area=default)  

- ### schemeType  
 **Type**: string  
 **Description**: [DataCite schemeType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=schemeType&check_keywords=yes&area=default)  

- ### resourceTypeGeneral  
 **Type**: string  
 **Description**: [DataCite resourceTypeGeneral](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=resourceTypeGeneral&check_keywords=yes&area=default)  

- **allOf**: #/definitions/relatedObject  
- **unevaluatedProperties**: FALSE  
- **properties**: string, #/definitions/relatedIdentifierType  
- **required**: relatedIdentifier, relatedIdentifierType, relationType  
- **if**: #/definitions/relatedObjectIf  
- **else**: #/definitions/relatedObjectElse   </details>  
 
