---
title: "Schema Overview"
output: 
  rmarkdown::html_vignette:
    css: styles.css
vignette: >
  %\VignetteIndexEntry{Schema Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Summary 

This document provides background information about data standards, json-schemas in general and the structure of the Wildlife Disease Data standard specifically.


### What is a JSON-schema

A JSON-schema is a human and machine readable document that defines a data standard by describing the structure, properties, and constraints for a dataset.
For those of us more accustomed to thinking about spreadsheet files and data frames, a property is roughly equivalent to a field or column. 
The JSON-schema defines the rules around the type of data used in  particular property (character, numeric, logical, etc), and its values (e.g. `massUnits` must be one of kg, mg, or g; `latitude` must be between -90 and 90; `sampleID` must be unique). 
The schema also describes how those fields should be combined into a coherent whole (i.e. the structure of the dataset). 

In a JSON-schema, fields can have parent child relationships. A field may itself be schema. For example, the `data` property in this
standard defines a `data` object that is a flat table with constraints, types, and/or
requirements. 
In this way, JSON-schema allows for the construction of modular 
schema documents that can leverage existing schemas (e.g. darwin core, or datacite).

Once we have created a schema, we can then validate data against it. 
The validation process happens via a validation engine and tells us if the data conform to the standard.
If the data do not conform, then the validation engine tells us precisely where the data are non-conformant and what the data standard expected to see. 

![](https://json-schema.org/img/json_schema.svg){width=95%}

For more detailed information see [JSON-Schema.org](https://json-schema.org/overview/what-is-jsonschema)

## Wildlife Disease Data Standard (WDDS) Structure

The Wildlife Disease Data Standard is composed of two sub-schemas (1) `disease_data` and
(2) `project_metadata`. 

`disease_data` describes the structure and contents of the wildlife disease data. 
It has certain required fields and is extensible. 
This data should be stored as a [tidy](https://doi.org/10.18637/jss.v059.i10) dataset in a flat file like a CSV.
This component of the standard relies heavily on the [Darwin Core data standard](https://dwc.tdwg.org/).

`project_metadata` describes the structure and contents of the descriptive 
metadata. That is, metadata about the project that enables discovery, 
identification, and attribution. This component of the standard relies heavily
on the [Data Cite Metadata Schema](https://datacite-metadata-schema.readthedocs.io/en/4.5/).

Researchers may validate their data against each sub-schema separately, or
use them in tandem to validate an entire data package. The term "data package" 
refers to a list or JSON object that contains both the `disease_data` and 
`project_metadata` components.

## Important vocabulary

**Property**: synonymous with field or column in a table. A property corresponds to a particular attribute (e.g. age, collectedBy, latitude, etc) of the data.  
**Required**: A property is must be included for a given schema or object within a schema.  
**Type**: [Type of data](https://json-schema.org/understanding-json-schema/reference/type). Common values include array, object, string, number, integer, null, and boolean.   
**Array**: A comma separated group of values. Similar to a vector in R but a little more flexible.  
**Array Items**: Array items define acceptable values for an array.  
  - minItems - how many items must be present in the array
  - minimum - inclusive - smallest value allowed in an array
  - maximum - inclusive - largest value allowed in an array
  - enum - controlled vocabulary for an array


## Terms

Below is a list of terms for the data standard. 
It is created directly from the `wdds_schema.json`
file and reflects the structure of that file. 
`disease_data` and `project_metadata` are top level properties of the schema, with many sub-properties.
Any item marked as <span style="color: black;background-color: #ffc107;">REQUIRED</span> must be included. 

```{r terms, eval=FALSE, include=FALSE, results='asis'}

wddsWizard::schema_terms |>
  cat()
```

## disease_data  
**Type**: object  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> Wildlife disease data. Stored in tidy form.  
**Required Fields**: sampleID, latitude, longitude, sampleCollectionMethod, hostIdentification, detectionTarget, detectionMethod, detectionOutcome, parasiteIdentification  
 **Reference**: schemas/disease_data.json  

- ### sampleID  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> A researcher-generated unique ID for the sample: usually a unique string of both characters and integers (e.g., OS BZ19-114 to indicate an oral swab taken from animal BZ19-114; see worked example below), to avoid conflicts that can arise when datasets are merged with number-only notation for samples. Ideally, sample names should be kept consistent across all online databases and physical resources (e.g., museum collections or project-specific sample archives).   
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### animalID  
**Type**: array  
**Description**: A researcher-generated unique ID for the individual animal from which the sample was collected: usually a unique string of both characters and integers (e.g., BZ19-114 to indicate animal 114 sampled in 2019 in Belize). Ideally, animal names should again be kept consistent across online databases and physical resources.   
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### latitude  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> Latitude of the collection site in decimal format. See http://rs.tdwg.org/dwc/terms/decimalLatitude   
**Array Items**
    - **type**: number, null  
    - **minItems**: 1  
    - **maximum**: 90  
    - **minimum**: -90    

- ### longitude  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> Longitude of the collection site  in decimal format. See http://rs.tdwg.org/dwc/terms/decimalLongitude   
**Array Items**
    - **type**: number, null  
    - **minItems**: 1  
    - **maximum**: 180  
    - **minimum**: -180    

- ### spatialUncertainty  
**Type**: array  
**Description**: Coordinate uncertainty from GPS recordings, post-hoc digitization, or systematic alterations (e.g., jittering or rounding) expressed in meters. See http://rs.tdwg.org/dwc/terms/coordinateUncertaintyInMeters   
**Array Items**
    - **type**: number, null  
    - **minItems**: 1  
    - **minimum**: 0    

- ### collectionDay  
**Type**: array  
**Description**: The day of the month on which the specimen was collected. See http://rs.tdwg.org/dwc/terms/day   
**Array Items**
    - **type**: integer, null  
    - **minItems**: 1  
    - **minimum**: 1  
    - **maximum**: 31    

- ### collectionMonth  
**Type**: array  
**Description**: The month in which the specimen was collected. See http://rs.tdwg.org/dwc/terms/month   
**Array Items**
    - **type**: integer, null  
    - **minItems**: 1  
    - **minimum**: 1  
    - **maximum**: 12    

- ### collectionYear  
**Type**: array  
**Description**: The year in which the specimen was collected. See http://rs.tdwg.org/dwc/terms/year   
**Array Items**
    - **type**: integer, null    

- ### sampleCollectionMethod  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> The technique used to acquire the sample and/or the tissue from which the sample was acquired (e.g. visual inspection; swab; wing punch; necropsy).  
**Example Values**: visual inspection, swab, wing punch, necropsy     
**Array Items**
    - **type**: string  
    - **minItems**: 1    

- ### sampleMaterial  
**Type**: array  
**Description**: Organic tissue or fluid being collected (e.g., “liver”; “blood”; “skin”; “whole organism”).  
**Example Values**: liver, blood, skin, whole organism     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### sampleCollectionBodyPart  
**Type**: array  
**Description**: Part of the animal body that samples are generated or collected from (e.g., “rectum”; “wing”).  
**Example Values**: rectum, wing     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### hostIdentification  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> The Linnaean classification of the animal from which the sample was collected, reported at the lowest possible level (ideally, species binomial name: e.g., Odocoileus virginianus or Ixodes scapularis). As necessary, researchers may also include an additional field indicating when uncertainty exists in the identification of the host organism (see Adding new fields). See http://rs.tdwg.org/dwc/terms/scientificName   
**Array Items**
    - **type**: string, null  
    - **minItems**: 1  
    - **not**: [HOMOhomo]{4} [SAPIENSsapiens]{7}    

- ### organismSex  
**Type**: array  
**Description**: The sex of the individual animal from which the sample was collected. See http://rs.tdwg.org/dwc/terms/sex  
**Example Values**: male, female, hermaphrodite     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### liveCapture  
**Type**: array  
**Description**: Whether the individual animal from which the sample was collected was alive at the time of capture. Should be TRUE or FALSE; lethal sampling should be recorded as TRUE as this field describes the organism at the time of capture.   
**Array Items**
    - **type**: boolean, null  
    - **minItems**: 1    

- ### hostLifeStage  
**Type**: array  
**Description**: The life stage of the animal from which the sample was collected (as appropriate for the organism) (e.g., juvenile, adult). See http://rs.tdwg.org/dwc/terms/lifeStage  
**Example Values**: juvenile, adult, larva     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### age  
**Type**: array  
**Description**: The numeric age of the animal from which the sample was collected, at the time of sample collection, if known (e.g., in monitored populations).   
**Array Items**
    - **type**: number, null  
    - **minItems**: 1  
    - **minimum**: 0    

- ### ageUnits  
**Type**: array  
**Description**: The units in which age is measured (usually years).   
**Array Items**
    - **type**: string, null  
    - **enum**: years, months, days, hours, minutes, seconds  
    - **minItems**: 1    

- ### mass  
**Type**: array  
**Description**: The mass of the animal from which the sample was collected, at the time of sample collection.   
**Array Items**
    - **type**: number, null  
    - **minItems**: 1  
    - **minimum**: 0    

- ### massUnits  
**Type**: array  
**Description**: The units that mass is recorded in (e.g., kg).  
**Example Values**: kg, g, mg, kilogram, milligram     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### length  
**Type**: array  
**Description**: The numeric length of the animal from which the sample was collected, at the time of sample collection.   
**Array Items**
    - **type**: number, null  
    - **minItems**: 1  
    - **minimum**: 0    

- ### lengthMeasurement  
**Type**: array  
**Description**: The axis of measurement for the organism being measured (e.g., snout-vent length or just SVL; wing length; primary feather).  
**Example Values**: snout-vent length, intertegular distance, primary feather     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### lengthUnits  
**Type**: array  
**Description**: The units that length is recorded in (e.g., meters).  
**Example Values**: mm, meters, cm, km     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### organismQuantity  
**Type**: array  
**Description**: A number or enumeration value for the quantity of organisms. See http://rs.tdwg.org/dwc/terms/organismQuantity  
**Example Values**: 1, 1.4, 12     
**Array Items**
    - **type**: number, null  
    - **minItems**: 1  
    - **minimum**: 0    

- ### organismQuantityUnits  
**Type**: array  
**Description**: The units that organism quantity is recorded in (e.g. “individuals”). See http://rs.tdwg.org/dwc/iri/organismQuantityType  
**Example Values**: individual, biomass, Braun-Blanquet scale     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### detectionTarget  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> The taxonomic identity of the parasite being screened for in the sample. This will often be coarser than the identity of a specific parasite identified in the sample: for example, in a study screening for novel bat coronaviruses, the entire family Coronaviridae might be the target; in a parasite dissection, the targets might be Acanthocephala, Cestoda, Nematoda, and Trematoda. For deep sequencing approaches (e.g., metagenomic and metatranscriptomic viral discovery), researchers should report each alignment target used as a new test to maximize reporting of negative data, or alternatively, select a subset that reflect specific study objectives and the focus of analysis (e.g., specific viral families). See http://rs.tdwg.org/dwc/terms/associatedOccurrences   
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### detectionMethod  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> The type of test performed to detect the parasite or parasite-specific antibody (e.g., 'qPCR', ‘ELISA’)   
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### forwardPrimerSequence  
**Type**: array  
**Description**: The sequence of the forward primer used for parasite detection (e.g., for a pan-coronavirus primer: 5’ CDCAYGARTTYTGYTCNCARC 3'). (Strongly encouraged if applicable, e.g., for PCR.)  
**Example Values**: 5’ CDCAYGARTTYTGYTCNCARC 3'     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### reversePrimerSequence  
**Type**: array  
**Description**: The sequence of the reverse primer used for parasite detection (e.g., 5’ RHGGRTANGCRTCWATDGC 3'). (Strongly encouraged if applicable, e.g., for PCR.)  
**Example Values**: 5’ RHGGRTANGCRTCWATDGC 3'     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### geneTarget  
**Type**: array  
**Description**: The parasite gene targeted by the primer (e.g. “RdRp” for PCR.).  
**Example Values**: RdRp     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### primerCitation  
**Type**: array  
**Description**: Citation(s) for the primer(s) (ideally doi, or other permanent identifier for a work, e.g. PMID).   
**Example Values**: https://doi.org/10.1016/j.virol.2007.06.009, Complete genome sequence of bat coronavirus HKU2 from Chinese horseshoe bats revealed a much smaller spike gene with a different evolutionary lineage from the rest of the genome, PMC7103351, https://openalex.org/works/w2036144053     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### probeTarget  
**Type**: array  
**Description**: Antibody or antigen targeted for detection. (Strongly encouraged if applicable, e.g., for ELISA.)   
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### probeType  
**Type**: array  
**Description**: Antibody or antigen used for detection. (Strongly encouraged if applicable, e.g., for ELISA.)   
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### probeCitation  
**Type**: array  
**Description**: Citation(s) for the probe(s) (ideally doi, or other permanent identifier for a work, e.g. PMID).    
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### detectionOutcome  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> The test result (i.e., positive, negative, or inconclusive). To avoid ambiguity, these specific values are suggested over numeric values (0 or 1). See http://rs.tdwg.org/dwc/terms/occurrenceStatus   
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### detectionMeasurement  
**Type**: array  
**Description**: Any numeric measurement of parasite detection that is more detailed than simple positive or negative results (e.g., viral titer, parasite counts, sequence reads).   
**Array Items**
    - **type**: number, null  
    - **minItems**: 1    

- ### detectionMeasurementUnits  
**Type**: array  
**Description**: Units for quantitative measurements of parasite intensity or test results (e.g., Ct, TCID50/mL, or parasite count).  
**Example Values**: Ct, TCID50/mL, parasite count     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### parasiteIdentification  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> The identity of a parasite detected by the test, if any, reported to the lowest possible taxonomic level, either as a Linnaean binomial classification or within the convention of a relevant taxonomic authority (e.g., Borrelia burgdorferi or Zika virus). Parasite identification may be more specific than detection target.  
**Example Values**: Zika virus, Borrelia burgdorferi, Onchocerca volvulus     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### parasiteID  
**Type**: array  
**Description**: A researcher-generated unique ID for an individual parasite (primarily useful in nested cases where this ID is used as an animal ID in another row, such as pathogen  testing of a blood-feeding arthropod removed from a vertebrate host).  
**Example Values**: 001, TICK201923     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### parasiteLifeStage  
**Type**: array  
**Description**: The life stage of the parasite from which the sample was collected (as appropriate for the organism) (e.g., juvenile, adult).  
**Example Values**: juvenile, adult, sporozoite     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    

- ### genbankAccession  
**Type**: array  
**Description**: The GenBank accession for any parasite genetic sequence(s), if appropriate.  Accession numbers or other identifiers for related data stored on another platform should be added in a different field (e.g. GISAID Accession, Immport Accession). See http://rs.tdwg.org/dwc/terms/otherCatalogNumbers   
**Example Values**: U49845 | U49846, U11111     
**Array Items**
    - **type**: string, null  
    - **minItems**: 1    
  
 

## project_metadata  
**Type**: object  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> Metadata for a project that largely follows the Datacite data standard.  
**Required Fields**: methodology, creators, titles, publicationYear, language, descriptions, fundingReferences  
 **Reference**: schemas/project_metadata.json  

- ### methodology  
**Type**: object  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> A broad categorization of how data were collected.  
**Properties**:  

    - ### eventBased  
**Type**: boolean  
**Description**: Whether or not research was conducted in response to a known or suspected infectious disease outbreak, observed animal morbidity or mortality, etc.  
    - ### archival  
**Type**: boolean  
**Description**: Whether samples were from an archival source (e.g., museum collections, biobanks).  
  

- ### creators  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> The full names of the creators. Should be in the format familyName, givenName.   
**Array Items**
    - ### name  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite name](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=name&check_keywords=yes&area=default)  

    - ### nameType  
**Type**: string  
**Description**: [DataCite nameType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=nameType&check_keywords=yes&area=default)  

    - ### givenName  
**Type**: string  
**Description**: [DataCite givenName](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=givenName&check_keywords=yes&area=default)  

    - ### familyName  
**Type**: string  
**Description**: [DataCite familyName](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=familyName&check_keywords=yes&area=default)  

    - ### nameIdentifiers  
**Type**: array  
**Description**: [DataCite nameIdentifiers](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=nameIdentifiers&check_keywords=yes&area=default)   
**Array Items**
        - ### nameIdentifier  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite nameIdentifier](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=nameIdentifier&check_keywords=yes&area=default)  
        - ### nameIdentifierScheme  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite nameIdentifierScheme](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=nameIdentifierScheme&check_keywords=yes&area=default)  
        - ### schemeUri  
**Type**: string  
**Description**: [DataCite schemeUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=schemeUri&check_keywords=yes&area=default)  
  

    - ### affiliation  
**Type**: array  
**Description**: [DataCite affiliation](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=affiliation&check_keywords=yes&area=default)   
**Array Items**
        - ### name  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite name](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=name&check_keywords=yes&area=default)  
        - ### affiliationIdentifier  
**Type**: string  
**Description**: [DataCite affiliationIdentifier](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=affiliationIdentifier&check_keywords=yes&area=default)  
        - ### affiliationIdentifierScheme  
**Type**: string  
**Description**: [DataCite affiliationIdentifierScheme](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=affiliationIdentifierScheme&check_keywords=yes&area=default)  
        - ### schemeUri  
**Type**: string  
**Description**: [DataCite schemeUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=schemeUri&check_keywords=yes&area=default)  
  

    - ### lang  
**Type**: string  
**Description**: [DataCite lang](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=lang&check_keywords=yes&area=default)  
  

- ### titles  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> A name or title by which a resource is known.   
**Array Items**
    - ### title  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite title](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=title&check_keywords=yes&area=default)  
    - ### titleType  
**Type**: string  
**Description**: [DataCite titleType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=titleType&check_keywords=yes&area=default)  
    - ### lang  
**Type**: string  
**Description**: [DataCite lang](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=lang&check_keywords=yes&area=default)  
  

- ### identifier  
**Type**: array  
**Description**: A unique string that identifies a resource.   
**Array Items**
    - ### identifier  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite identifier](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=identifier&check_keywords=yes&area=default)  
    - ### identifierType  
**Type**: string  
**Description**: [DataCite identifierType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=identifierType&check_keywords=yes&area=default)  
  

- ### subjects  
**Type**: array  
**Description**: Subject, keyword, classification code, or key phrase describing the resource.   
**Array Items**
    - ### subject  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite subject](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=subject&check_keywords=yes&area=default)  
    - ### subjectScheme  
**Type**: string  
**Description**: [DataCite subjectScheme](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=subjectScheme&check_keywords=yes&area=default)  
    - ### schemeUri  
**Type**: string  
**Description**: [DataCite schemeUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=schemeUri&check_keywords=yes&area=default)  
    - ### valueUri  
**Type**: string  
**Description**: [DataCite valueUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=valueUri&check_keywords=yes&area=default)  
    - ### classificationCode  
**Type**: string  
**Description**: [DataCite classificationCode](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=classificationCode&check_keywords=yes&area=default)  
    - ### lang  
**Type**: string  
**Description**: [DataCite lang](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=lang&check_keywords=yes&area=default)  
  

- ### publicationYear  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> The year when the data was or will be made publicly available.  

- ### rights  
**Type**: array  
**Description**: Any rights information for this resource.   
**Array Items**
    - ### rights  
**Type**: string  
**Description**: [DataCite rights](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=rights&check_keywords=yes&area=default)  
    - ### rightsUri  
**Type**: string  
**Description**: [DataCite rightsUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=rightsUri&check_keywords=yes&area=default)  
    - ### rightsIdentifier  
**Type**: string  
**Description**: [DataCite rightsIdentifier](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=rightsIdentifier&check_keywords=yes&area=default)  
    - ### rightsIdentifierScheme  
**Type**: string  
**Description**: [DataCite rightsIdentifierScheme](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=rightsIdentifierScheme&check_keywords=yes&area=default)  
    - ### schemeUri  
**Type**: string  
**Description**: [DataCite schemeUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=schemeUri&check_keywords=yes&area=default)  
    - ### lang  
**Type**: string  
**Description**: [DataCite lang](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=lang&check_keywords=yes&area=default)  
  

- ### descriptions  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> All additional information that does not fit in any of the other categories. May be used for technical information or detailed information associated with a scientific instrument.   
**Array Items**
    - ### description  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite description](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=description&check_keywords=yes&area=default)  
    - ### descriptionType  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite descriptionType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=descriptionType&check_keywords=yes&area=default)  
    - ### lang  
**Type**: string  
**Description**: [DataCite lang](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=lang&check_keywords=yes&area=default)  
  

- ### language  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> The primary language of the resource.  

- ### fundingReferences  
**Type**: array  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> Name and other identifying information of a funding provider.   
**Array Items**
    - ### funderName  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite funderName](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=funderName&check_keywords=yes&area=default)  
    - ### funderIdentifier  
**Type**: string  
**Description**: [DataCite funderIdentifier](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=funderIdentifier&check_keywords=yes&area=default)  
    - ### funderIdentifierType  
**Type**: string  
**Description**: [DataCite funderIdentifierType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=funderIdentifierType&check_keywords=yes&area=default)  
    - ### awardNumber  
**Type**: string  
**Description**: [DataCite awardNumber](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=awardNumber&check_keywords=yes&area=default)  
    - ### awardUri  
**Type**: string  
**Description**: [DataCite awardUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=awardUri&check_keywords=yes&area=default)  
    - ### awardTitle  
**Type**: string  
**Description**: [DataCite awardTitle](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=awardTitle&check_keywords=yes&area=default)  
  

- ### relatedIdentifiers  
**Type**: array  
**Description**: [DataCite relatedIdentifiers](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=relatedIdentifiers&check_keywords=yes&area=default)   
**Array Items**
    - ### relationType  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite relationType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=relationType&check_keywords=yes&area=default)  

    - ### relatedMetadataScheme  
**Type**: string  
**Description**: [DataCite relatedMetadataScheme](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=relatedMetadataScheme&check_keywords=yes&area=default)  

    - ### schemeUri  
**Type**: string  
**Description**: [DataCite schemeUri](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=schemeUri&check_keywords=yes&area=default)  

    - ### schemeType  
**Type**: string  
**Description**: [DataCite schemeType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=schemeType&check_keywords=yes&area=default)  

    - ### resourceTypeGeneral  
**Type**: string  
**Description**: [DataCite resourceTypeGeneral](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=resourceTypeGeneral&check_keywords=yes&area=default)  
  
    - ### relatedIdentifier  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite relatedIdentifier](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=relatedIdentifier&check_keywords=yes&area=default)  

    - ### relatedIdentifierType  
**Type**: string  
**Description**: <span style="color: black;background-color: #ffc107;">REQUIRED</span> [DataCite relatedIdentifierType](https://datacite-metadata-schema.readthedocs.io/en/4.5/search/?q=relatedIdentifierType&check_keywords=yes&area=default)  
  
  
 
