---
title: "Schema Overview"
output: 
  rmarkdown::html_vignette:
    css: styles.css
vignette: >
  %\VignetteIndexEntry{Schema Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Summary 

This document provides background information about data standards, json-schemas in general and the structure of the Wildlife Disease Data standard specifically.

### Why do we need a data standard?

Data standards facilitate the sharing, (re)use, and aggregation of data by
humans and machines through the use of a common structure, set of properties, and vocabulary. 

Using a data standard makes it easier to share datasets and reproduce results. Data standards facilitate sharing by lowering the effort it takes to document the dataset. 
Sharing data makes it easier to reproduce results. 
Sharing data is especially important in disciplines like disease ecology where producing datasets is both resource intensive and limited, making it unlikely that someone will be able to replicate field work for the sake of verifying a finding.
Being able to reproduce results increases confidence in findings and allows others (most likely the original dataset producers) to build on those results. 

Compliance with a data standard eases data interpretation. 
Because the data have been validated against the standard, a researcher knows that fields in the dataset meet set definitions and the data conform to a certain structure. 
Researchers can also identify deviations from the standard more readily because of the descriptions and examples provided in the standard.

Compliant datasets can be aggregated with other datasets that meet the data standard.
As the number of datasets that use the standard grows, the questions that researchers can ask of the data changes.
For example, if we have datasets describing disease incidence for some pathogen from dozens of locations across multiple regions, we can explore general modeling frameworks for that pathogen using a unified dataset. 

By using a data standard we can contribute to a virtuous cycle that promotes high quality open and reproducible science through sharing, reuse, and aggregation of datasets. 

### What is a JSON-schema

A JSON-schema is a human and machine readable document that defines a data standard by describing the structure, properties, and constraints for a dataset.
For those of us more accustomed to thinking about spreadsheet files and data frames, a property is roughly equivalent to a field or column. 
The JSON-schema defines the rules around the type of data used in  particular property (character, numeric, logical, etc), and its values (e.g. `massUnits` must be one of kg, mg, or g; `latitude` must be between -90 and 90; `sampleID` must be unique). 
The schema also describes how those fields should be combined into a coherent whole (i.e. the structure of the dataset). 

In a JSON-schema, fields can have parent child relationships. A field may itself be schema. For example, the `data` property in this
standard defines a `data` object that is a flat table with constraints, types, and/or
requirements. 
In this way, JSON-schema allows for the construction of modular 
schema documents that can leverage existing schemas (e.g. dawrin core, or datacite).

Once we have created a schema, we can then validate data against it. 
The validation process happens via a validation engine and tells us if the data conform to the standard.
If the data do not conform, then the validation engine tells us precisely where the data are non-conformant and what the data standard expected to see. 

![](https://json-schema.org/img/json_schema.svg){width=95%}

For more detailed information see [JSON-Schema.org](https://json-schema.org/overview/what-is-jsonschema)

## How do I use this package and the Wildlife Disease Data Standard?

This package can be used to explore Wildlife Disease Data Standard and validate data. 

See our preprint [A minimum data standard for wildlife disease studies](https://doi.org/10.32942/X2TW4J) for discussions of appropriate use and  more complex data structures (e.g. pooled samples, parasites with an arthropod vector, etc.).


### Starting from scratch - don't have any data, yet!

We strongly recommend creating a Data Management Plan (DMP) at the beginning of any project and seriously considering the data life cycle for your project.

Use our CSV templates and the data standard as a guide for storing data. 
Make sure that at least the minimum required fields are captured in data collection, and that they match the data standard.
Lab and field data may not immediately come out in a "tidy" format but as long as all properties are captured and data remain disaggregated (1 row = 1 observation), it should be possible to reshape the data to be compatible with the standard.


### Starting from an existing project - I've got data!?

This is potentially more challenging than starting from scratch and requires harmonizing your data and the data standard.
First, check that your data are or can be disaggregated to a 1 row = 1 observation tidy data model. 
Then, cross-walk the fields from your data with the properties in the standard and make sure that your data contain all the required fields. 
Next, start to reshape it into a single, tidy, table.
You will need to re-name fields that correspond to the those in the standard. 

### Validating your data

This package has vignettes describing how to move from CSV/excel to JSON for
[project metadata](project_metadata.html) and [wildlife disease data](disease_data.html).


## Wildlife Disease Data Standard (WDDS) Structure

The Wildlife Disease Data Standard is composed of two sub-schemas (1) `disease_data` and
(2) `project_metadata`. 

`disease_data` describes the structure and contents of the wildlife disease data. 
It has certain required fields and is extensible. 
This data should be stored as a [tidy](https://doi.org/10.18637/jss.v059.i10) dataset in a flat file like a CSV.
This component of the standard relies heavily on the [Darwin Core data standard](https://dwc.tdwg.org/).

`project_metadata` describes the structure and contents of the descriptive 
metadata. That is, metadata about the project that enables discovery, 
identification, and attribution. This component of the standard relies heavily
on the [Data Cite Metadata Schema](https://datacite-metadata-schema.readthedocs.io/en/4.5/).

Researchers may validate their data against each sub-schema separately, or
use them in tandem to validate an entire data package. The term "data package" 
refers to a list or JSON object that contains both the `disease_data` and 
`project_metadata` components.

## Important vocabulary

**Property**: synonymous with field or column in a table. 
**Required**: A property is must be included for a given schema or object within a schema.
**Array**: A comma separated group of values. Similar to a vector in R but a little more flexible. 


# Schema terms

Below is a list of schema terms. 
It is created directly from the `wdds_schema.json`
file and reflects the structure of that file. 
`disease_data` and `project_metadata` are top level properties of the schema, with many sub-properties.
Any item marked as <req> REQUIRED </req> must be included. 

```{r terms, eval=TRUE, include=FALSE, results="asis"}
 library(wddsWizard)
create_schema_docs() |>
  cat()
```

