---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wddsWizard)
library(jsonvalidate)
```

This package is home to the [Wildlife Disease Data Standard](https://doi.org/10.32942/X2TW4J).
It provides template csv files for storing data and a set of functions for validating datasets against the standard. 

## The Whole game

Flat data files (csv/xlsx) are imported into R, lightly restructured, and then converted to JSON and validated against the Wildlife Disease Data Standard. 
Data sets either pass (ðŸŽ‰) or fail with informative errors. 

## Getting Started

The data standard is designed to be flexible and accessible. 
It is composed of two components: disease data and project metadata.
The disease data component describes the contents and structure of data related to the detection (or not) of a parasite in a given host. 
The project metadata component describes the contents and structure of data related to the creation of the disease data component. 
The disease data component allows us to create a collection of datasets that can be re-used, aggregated, and shared, while the project metadata component provides context for the data, makes it easier to find the dataset, and gives clear information about attribution and use (rights/license). 

Using this package, you can validate one or both components of the data standard. 
It will be helpful to open the [Data Standard Terms](schema_overview.html#terms) in a separate browser tab as you go through the vignette and explore templates. 

If you use the templates provided, you can focus on the field descriptions and largely ignore types and array items.

### Templates

We have created several templates to make it easier to get started using the Wildlife Disease Data Standard.

```{r templates}
# list templates
use_wdds_template()
```

You can make your own copies of the template files with the `use_wdds_template` function.

```{r use-template, eval=FALSE}
use_wdds_template("disease_data_template.csv", file_name = "my_interesting_disease_data.csv", open = TRUE)

use_wdds_template("project_metadata_template.csv", file_name = "my_project_metadata.csv", open = TRUE)
```

The templates may contain additional fields that are not strictly required. Required fields are clearly marked in [Data Standard Terms](schema_overview.html#terms) documentation.

The following fields are required for disease data:
```{r required-fields-dd}
wddsWizard::disease_data_required_fields
```

The following fields are required for project metadata:
```{r required-fields-pmd}
wddsWizard::project_metadata_required_fields
```

### Disease Data

Disease data are expected to be in a "tidy" form (think CSV or XLS where each row is an observation and each column is a property). 
Each column in the table that is part of the standard will be validated. 
You may include additional columns as needed. 

Disease data and project metadata can be validated separately. 
In the code below we will read in a csv file, do some light wrangling, transform the data to JSON, and then validate the data. 

```{r disease-data, warning=FALSE}
## read in the data
my_disease_data <- wdds_example_data(version = "latest", file = "my_interesting_disease_data.csv") |>
  read.csv()

# clean up field names to match JSON schema
my_disease_data <- clean_field_names(my_disease_data)

### Check for required Fields -

# check that all required fields are in the data
all(wddsWizard::disease_data_required_fields %in% names(my_disease_data))

## Prep for JSON

my_disease_data_prepped <- prep_data(my_disease_data)

## make the JSON!
my_disease_data_json <- my_disease_data_prepped |>
  jsonlite::toJSON(pretty = TRUE)

### validate the JSON

# get the schema file
schema <- wdds_json(version = "latest", file = "schemas/disease_data.json")

# this creates a function that we can use to validate our data
dd_validator <- jsonvalidate::json_validator(schema, engine = "ajv")

# use the validator to check if the disease data conforms to the disease_data component of the standard
dd_validation <- dd_validator(my_disease_data_json, verbose = TRUE)

## check for errors!

errors <- attributes(dd_validation)

if (!dd_validation) {
  errors$errors
} else {
  print("Valid disease data!ðŸ˜")
}
```

### Project Metadata

Project metadata largely follow the [Datacite Metadata Schema](https://datacite-metadata-schema.readthedocs.io/en/4.5/). 
Again, the data standard allows you to include additional properties.

Note that if you are comfortable with JSON, it may be easier to write project metadata directly as JSON. 

In the example below, we will use project metadata created from the  project metadata template to create JSON that can be validated against the project metadata component of the data standard.  

```{r prep-metadata, warning=FALSE}
# read in project metadata created from template

my_project_metadata <- wdds_example_data(version = "latest", file = "my_project_metadata.csv") |>
  read.csv()

# prepare project metadata
my_project_metadata_prepped <- prep_from_metadata_template(my_project_metadata)


# check that all required fields are in the project metadata
all(wddsWizard::project_metadata_required_fields %in% names(my_project_metadata_prepped))

# convert to json

my_project_metadata_json <- my_project_metadata_prepped |>
  jsonlite::toJSON(pretty = TRUE)

# validate against project metadata schema

schema <- wdds_json(version = "latest", file = "schemas/project_metadata.json")

pm_validator <- jsonvalidate::json_validator(schema, engine = "ajv")

pm_validation <- pm_validator(my_project_metadata_json, verbose = TRUE)

## check for errors!

errors <- attributes(pm_validation)

if (!pm_validation) {
  errors$errors
} else {
  print("Valid project metadata!ðŸ˜")
}
```

See the vignettes on [Project Metadata](project_metadata.html) and [Wildlife Disease Data](disease_data.html) for more
details on preparing those components. 


## Combine disease data and project metadata

Finally we will check the disease data and the project metadata against the standard.

### Combine components

The first thing we have to do is combine the disease data and project metadata components in a list and check that we have all the required fields. 

```{r combine-data}
## use append so that you do not add levels to your list

data_package <- list(
  disease_data = my_disease_data_prepped,
  project_metadata = my_project_metadata_prepped
)

# check that all required fields are in the data

req_field_check <- wddsWizard::schema_required_fields %in% names(data_package)

if (all(!req_field_check)) {
  wddsWizard::schema_required_fields[!req_field_check]
} else {
  print("all required fields present ðŸ¥³")
}
```

### Make JSON

Next we will convert the `data_package` so that it can be validated. 

```{r make-json}
# convert to json

data_package_json <- jsonlite::toJSON(data_package, pretty = TRUE)
```

### Validate your json!

Here we will use the {jsonvalidate} package to make sure `data_package_json` 
conforms to the wildlife disease data standard.

```{r json-validate, warning=FALSE}
schema <- wdds_json(version = "latest", file = "wdds_schema.json")

wdds_validator <- jsonvalidate::json_validator(schema, engine = "ajv")

project_validation <- wdds_validator(data_package_json, verbose = TRUE)

if (project_validation) {
  print("Your data package is valid! ðŸŽŠ ")
} else {
  errors <- attributes(project_validation)
  errors$errors
}
```

### Handling Errors

You're likely going to get some errors and interpretting them can sometimes be a challenge. 

Lets dive in by modifying the data_package to create a known issue. 

Looking at the data standard, we can see that `language` is  type: string. 

```{r proper-json-langauge}
# the properly formatted language property should like this:
data_package$project_metadata["language"]|>jsonlite::toJSON(pretty = TRUE)

```


```{r make-json-error}

# lets make language a multi-row data frame
data_package_error_lang <- data_package

data_package_error_lang$project_metadata$language <- data.frame(language = c("en","fr","cn"))

 jsonlite::toJSON(data_package_error_lang$project_metadata["language"], pretty = TRUE) 

```
Our new version of language is an array of objects - square braces`[]` = array, curly braces `{}` = object. So its definitely not of type string. So lets see what happens we when go to validate the metadata.

```{r validate-json-error}

data_package_error_lang_json <- jsonlite::toJSON(data_package_error_lang)

lang_issue <- wdds_validator(data_package_error_lang_json, verbose = TRUE)

errors_df <- attributes(lang_issue)$errors

errors_df

```
### Excuse me? What does this mean?

- **instancePath** = where did the schema find an error?
- **schemaPath** = where does the property come from? 
- **keyword** = what kind of violation occurred?
- **message** = what went wrong?
- **params** = what are the parameters for that property in the schema? 
- **schema** = what should be in that property?
- **parentSchema** = type and description from the schemaPath
- **data** = violating item you provided
- **dataPath** = where did the violation occur? 


SO to interpret our error its really helpful to start by looking at the `instancePath`, `keyword`, and `message`.

```{r initial_error_interpretation}

errors_df[c("instancePath","keyword","message")]

```

OH project metadata must be type string! 
You will notice that it does not tell you how it interpreted the object. 
You can figure this out by converting `errors_df$data` to json. 


```{r deeper_look_error}
errors_df$data[[1]] |> jsonlite::toJSON(pretty = TRUE)

```
This is obviously not a string!

### How do I correct this error? 

Since our work is multi-lingual, we might think that `c("en","fr","cn")` would be appropriate but that would be converted to an `array` in JSON. 

lets try just `"en, fr, cn"` and see if that works

```{r correct-lang-error}
data_package_correct_lang <- data_package_error_lang

data_package_correct_lang$project_metadata$language <- "en, fr, cn"

data_package_correct_lang_json <- jsonlite::toJSON(data_package_correct_lang, pretty = TRUE)

correct_lang_issue <- wdds_validator(data_package_correct_lang_json, verbose = TRUE)

if(!correct_lang_issue){
  print("WHAT! I gave it a string!?")
}

```

Yes you did! BUT if we take a peek at the json object we can see that your string
was converted to the array `"language": ["en, fr, cn"]`.

Why? because jsonlite errors on the side of wrapping things in arrays.

Thankfully the authors of this package have provided methods for preparing items!
Here we will use `prep_language` when we add the item. Prep_language is a thin
wrapper around `prep_atomic` - which ensures that jsonlite will interpret your
R object as numeric, string, boolean, or NULL.

```{r correct-lang-error-final}
data_package_correct_lang <- data_package_error_lang

data_package_correct_lang$project_metadata$language <- prep_language("en, fr, cn")


data_package_correct_lang_json <- jsonlite::toJSON(data_package_correct_lang, pretty = TRUE)

 wdds_validator(data_package_correct_lang_json, verbose = TRUE)


```


## Best practices for free text fields

We recommend that data producers use controlled vocabularies or ontologies when filling out free text fields. We recognize that selecting an appropriate vocabulary can be challenging and recommend the following platforms for finding appropriate terms.

Recommended ontology hosting and search platforms with distinct funding sources.
Name | URL
-- | --
Ontobee | https://ontobee.org/
Ontology Lookup Service | https://www.ebi.ac.uk/ols4/
BioPortal | https://bioportal.bioontology.org/


All three platforms allow users to search for terms stored in ontologies, explore relationships between terms, and find analogues. A user will have to explore a given ontology to find the most appropriate term. In Table S2 we list specific ontologies or authorities that may be appropriate for a given field.

Recommended ontologies or authorities for specific fields.
Field | URL
-- | --
Host Identification | https://www.gbif.org/species/search
Gene Target | https://www.ebi.ac.uk/ols4/ontologies/go
Sample Collection Method | http://purl.obolibrary.org/obo/OBI_0000659
Sample Collection Body Part | https://www.ebi.ac.uk/ols4/ontologies/uberon
Sample Collection Material | http://purl.obolibrary.org/obo/OBI_0001479




