---
title: "The Whole Game"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Whole Game}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(wddsWizard)
library(jsonvalidate)
```

This vignette explains how to put data and project metadata together into a 
single json file that can be validated against the data standard. 

See the vignettes on [Project Metadata](project_metadata.html) and [Wildlife Disease Data](disease_data.html) for more
details on preparing those components. The general strategy is to get each component
into a data frame, then use specific methods to ensure they are in the proper
form to be translated into json. 


## Combine disease data and project metadata

```{r combine-data}

## this is equivalent to the disease data from the vignette
becker_disease_data <- wddsWizard::becker_disease_data

## this is equivalent to the project metadata from the vignette

becker_project_metadata <- wddsWizard::becker_project_metadata

## use append so that you do not add levels to your list

data_package <- append(becker_disease_data,becker_project_metadata)

```

## Make JSON

```{r make-json}

# prepare for json - prepares data using known methods

data_package_prepped <- data_package |>
  wddsWizard::prep_for_json()

# check that all required fields are in the data

req_field_check <- wddsWizard::schema_required_fields %in% names(data_package_prepped)

if(all(!req_field_check)){
  wddsWizard::schema_required_fields[!req_field_check]  
}

# convert to json

data_package_json <- jsonlite::toJSON(data_package_prepped,pretty = TRUE)


```

## Validate your json!

Here we will use the {jsonvalidate} package to make sure our `data_package_json` 
meets the wildlife disease data standard.

```{r json-validate, warning=FALSE}

schema <- here::here("inst/extdata/wdds_schema/wdds_schema.json")

wdds_validator <- jsonvalidate::json_validator(schema,engine = "ajv")

project_validation <- wdds_validator(data_package_json,verbose = TRUE)

if(project_validation){
  print("Your data package is valid!")
} else {
errors <- attributes(project_validation)
errors$errors
}

## check for errors!


```

